---
title: "Stepwise LRT method"
subtitle: "2017 dataset"
author: "Frederick Restrepo Estrada"
date: "2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


Made in **R version 4.1.3 (2022-03-10)**


Data used: `snapshot2017you.sav`

The dataset is a trimmed version (snapshot) of a dataset made available by CBS. It is not publicly available.


Use this template as follows:

1. In the chunk under "Define", indicate if restrictions are considered. If models with restrictions are considered, comment out  `resTes <- ""`. If model with restrictions are **not** considered, comment out `resTes <- "R"` (`ctrl+shift+C`). 

2. Run all chunks (`ctrl+alt+r`).


Load libraries.
```{r}
library(brew)
library(haven)
```

Define functions.
```{r}
# function for extracting LG output
extract.output <- function(outPath, Nob) {
  
  output <- readLines(con = outPath) # read text lines
  output <- output[-1] # remove headers that are created when using `write` in LG
  output <- strsplit(x = output, split = ",") # LG creates `.csv` file
  maxLength <- max( sapply(X = output, FUN = length) ) # max number of columns in `.csv` file
  output <- lapply(X = output, FUN = `length<-`, maxLength) # create elements of equal length needed for dataframe
  output <- as.data.frame( do.call(rbind, output) ) # create dataframe
  row.names(output) <- NULL
  
  nams <- unique( output[ , 1] )
  note <- output[ which( output[ , 4] == 0 ), 5 ]
  note[is.na(note)] <- "OK"
  
  LLs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 6 ] ) # log-likelihoods
  # LPs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 8 ] ) # log-posteriors
  # L2s <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 9 ] ) # likelihood-ratio chi-square statistics
  # BVRs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 13 ] ) # total bivariate residuals
  Npars <- as.numeric( rowSums( !is.na(output[ which( output[, 5] == "Parameters" ), 6:ncol(output) ] ) ) )
  AIC.LL <- -2*LLs + Npars*2
  BIC.LL <- -2*LLs + Npars*log(Nob)
  # BIC.L2 <- L2s - df*log(Nob)
  
  return( data.frame(nam = nams, LL = LLs, Npars = Npars, AIC.LL = AIC.LL, BIC.LL = BIC.LL, note = note) )
}

# function for comparing models with a LRT
likelihood.ratio.test <- function(LL1, LL2, Npar1, Npar2) {
  LR <- -2*(LL1 - LL2)
  DFdif <- Npar2 - Npar1
  pval <- pchisq(q = LR, df = DFdif, lower.tail = FALSE)
  return( list(LRTS = LR, df = DFdif,  p = pval) )
}
```


#### Stepwise LRT method for a single covariate Z

Define.
```{r}
# restriction 
# resTes <- ""
resTes <- "R"

covars <- c(
  "age_group",
  "company_size",
  "economic_activity",
  "education_level",
  "interview_style",
  "job_length",
  "migration_background",
  "software_cluster",
  "work_hours"
  )

dat <- read_sav(file = "snapshot2017you.sav")
N <- nrow(dat)
```

Step1-5.
```{r}
bestModels <- data.frame()

for ( i in 1:length(covars) ) {
  
  aTab <- table( dat[, "contract_ER"][[1]], dat[, covars[i]][[1]] )
  
  if ( resTes == ""  ) {
    restrict <- "without"
  } else { # R7 and R8
    restrict <- ifelse(test = any(aTab == 0), yes = "with", no = "without")
  }
  
  outNam1 <- paste0("output-", covars[i], "-2017.csv")
  outNam2 <- paste0("output-step2-", covars[i], "-2017.csv")
  
  a1 <- extract.output(outPath = outNam1, Nob = N) # LG output for step1, 3, 4, and 5
  a2 <- extract.output(outPath = outNam2, Nob = N) # LG output for step2
  
  ###########
  ## STEP1 ##
  ###########
  
  resultLRT <- likelihood.ratio.test(
    LL1 =   a1[ a1$nam == "ER: no DIF; LFS: no DIF"                  , "LL" ],
    LL2 =   a1[ a1$nam == "ER: non-uniform DIF; LFS: non-uniform DIF", "LL" ],
    Npar1 = a1[ a1$nam == "ER: no DIF; LFS: no DIF"                  , "Npars" ],
    Npar2 = a1[ a1$nam == "ER: non-uniform DIF; LFS: non-uniform DIF", "Npars" ] )
  
  if (resultLRT$p > 0.05) {
    ERdif2 <- ERdif4 <- "ER: no DIF"
    LFSdif2 <- LFSdif4 <- "LFS: no DIF"
  } else {
  
    ###########
    ## STEP2 ##
    ###########
      
    # M2.0.1 vs. M2.1.1
    resultLRT <- likelihood.ratio.test(
      LL1 =   a2[ a2$nam == "ER: no DIF"         , "LL" ], 
      LL2 =   a2[ a2$nam == "ER: non-uniform DIF", "LL" ],
      Npar1 = a2[ a2$nam == "ER: no DIF"         , "Npars" ],
      Npar2 = a2[ a2$nam == "ER: non-uniform DIF", "Npars" ] )
    # DIF ER (based on step2)
    ERdif2 <- ifelse( test = resultLRT$p > 0.05, yes = "ER: no DIF", no = "ER: non-uniform DIF" )
  
    # M2.0.2 vs. 2.1.2
    resultLRT <- likelihood.ratio.test(
      LL1 =   a2[ a2$nam == "LFS: no DIF"         , "LL" ], 
      LL2 =   a2[ a2$nam == "LFS: non-uniform DIF", "LL" ],
      Npar1 = a2[ a2$nam == "LFS: no DIF"         , "Npars" ],
      Npar2 = a2[ a2$nam == "LFS: non-uniform DIF", "Npars" ] )
    # DIF LFS (based on step2)
    LFSdif2 <- ifelse(test = resultLRT$p > 0.05, yes = "LFS: no DIF", no = "LFS: non-uniform DIF")
    
    if (paste0(ERdif2, "; ", LFSdif2) == "ER: no DIF; LFS: no DIF") {
      ERdif4 <- ERdif2
      LFSdif4 <- LFSdif2 
      print( paste0("failure step2; data: ", covars[i], "; M3.0 = M1.0") )
    } else {
      
      ###########
      ## STEP3 ##
      ###########
      
      # M1.0 vs. M3.0
      resultLRT <- likelihood.ratio.test(
        LL1 =   a1[ a1$nam == "ER: no DIF; LFS: no DIF",     "LL" ], # M1.0
        LL2 =   a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2), "LL" ], # M3.0
        Npar1 = a1[ a1$nam == "ER: no DIF; LFS: no DIF",     "Npars" ],
        Npar2 = a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2), "Npars" ] )
      if (resultLRT$p > 0.05) { print( paste0("failure step3; M1.0 vs. M3.0; data: ", covars[i]) ) }
      
      # M3.0 vs. M1.1
      resultLRT <- likelihood.ratio.test(
        LL1   = a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2),               "LL" ], # M3.0
        LL2   = a1[ a1$nam == "ER: non-uniform DIF; LFS: non-uniform DIF", "LL" ], # M1.1
        Npar1 = a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2),               "Npars" ],
        Npar2 = a1[ a1$nam == "ER: non-uniform DIF; LFS: non-uniform DIF", "Npars" ] )
      if (resultLRT$p <= 0.05) { print( paste0("failure step3; M3.0 vs. M1.1; data: ", covars[i]) ) }
    
      ###########
      ## STEP4 ##
      ###########
      
      # DIF ER (based on step4)
      # M4.1 vs. M3.0
      if (ERdif2 == "ER: no DIF") { ERdif4 <- "ER: no DIF"
      } else if ( ERdif2 == "ER: non-uniform DIF" ) { # if there is something to be tested
        resultLRT <- likelihood.ratio.test(
          LL1 =   a1[ a1$nam == paste0("ER: uniform DIF; ", LFSdif2), "LL" ], # M4.1
          LL2 =   a1[ a1$nam == paste0(ERdif2,        "; ", LFSdif2), "LL" ], # M3.0
          Npar1 = a1[ a1$nam == paste0("ER: uniform DIF; ", LFSdif2), "Npars" ],
          Npar2 = a1[ a1$nam == paste0(ERdif2,        "; ", LFSdif2), "Npars" ] )
        ERdif4 <- ifelse(test = resultLRT$p > 0.05, yes = "ER: uniform DIF", no = "ER: non-uniform DIF") 
      }
  
      # DIF LFS (based on step4)
      # M4.2 vs. M3.0
      if (LFSdif2 == "LFS: no DIF") { LFSdif4 <- "LFS: no DIF"
      } else if (LFSdif2 == "LFS: non-uniform DIF") { # if there is something to be tested
        resultLRT <- likelihood.ratio.test(
          LL1 =   a1[ a1$nam == paste0(ERdif2, "; LFS: uniform DIF" ), "LL" ], # M4.2
          LL2 =   a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2),         "LL" ], # M3.0
          Npar1 = a1[ a1$nam == paste0(ERdif2, "; LFS: uniform DIF" ), "Npars" ],
          Npar2 = a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2),         "Npars" ] )
        LFSdif4 <- ifelse(test = resultLRT$p > 0.05, yes = "LFS: uniform DIF", no = "LFS: non-uniform DIF" )
      }
      
      ###########
      ## STEP5 ##
      ###########
      
      # M5.0 vs. M3.0
      resultLRT <- likelihood.ratio.test(
        LL1 =   a1[ a1$nam == paste0(ERdif4, "; ", LFSdif4), "LL" ], # M5.0
        LL2 =   a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2), "LL" ], # M3.0
        Npar1 = a1[ a1$nam == paste0(ERdif4, "; ", LFSdif4), "Npars" ],
        Npar2 = a1[ a1$nam == paste0(ERdif2, "; ", LFSdif2), "Npars" ] )
      if (resultLRT$p <= 0.05) { print( paste0("failure step5; M5.0 vs. M3.0; data: ", covars[i]) ) }
      
      ############
      ## STEP5R ##
      ############
  
      if ( restrict == "with" ) {
        
      ## PART1 ##
        
      resultLRT <- likelihood.ratio.test(
        LL1 =   a1[ a1$nam == paste0( paste0(ERdif4, " (R7)"), "; ", LFSdif4), "LL" ], # M5.0R7
        LL2 =   a1[ a1$nam == paste0(        ERdif4,           "; ", LFSdif4), "LL" ], # M5.0
        Npar1 = a1[ a1$nam == paste0( paste0(ERdif4, " (R7)"), "; ", LFSdif4), "Npars" ],
        Npar2 = a1[ a1$nam == paste0(        ERdif4,           "; ", LFSdif4), "Npars" ] )
      if (resultLRT$p > 0.05) { ERdif5_R7 <- paste0(ERdif4, " (R7)") }

      resultLRT <- likelihood.ratio.test(
        LL1 =   a1[ a1$nam == paste0( paste0(ERdif4, " (R8)"), "; ", LFSdif4), "LL" ], # M5.0R8
        LL2 =   a1[ a1$nam == paste0(        ERdif4,           "; ", LFSdif4), "LL" ], # M5.0
        Npar1 = a1[ a1$nam == paste0( paste0(ERdif4, " (R8)"), "; ", LFSdif4), "Npars" ],
        Npar2 = a1[ a1$nam == paste0(        ERdif4,           "; ", LFSdif4), "Npars" ] )
      if (resultLRT$p > 0.05) { ERdif5_R8 <- paste0(ERdif4, " (R8)") }
        
      ## PART2 ##
        
      if ( exists(x = "ERdif5_R7") & exists(x = "ERdif5_R8") ) {
        resultLRT <- likelihood.ratio.test(
          LL1 =   a1[ a1$nam == paste("ER: no DIF;", LFSdif4), "LL" ], # M5.0R9
          LL2 =   a1[ a1$nam == paste0(ERdif4, "; ", LFSdif4), "LL" ], # M5.0
          Npar1 = a1[ a1$nam == paste("ER: no DIF;", LFSdif4), "Npars" ],
          Npar2 = a1[ a1$nam == paste0(ERdif4, "; ", LFSdif4), "Npars" ] )
        if (resultLRT$p > 0.05) {
          ERdif5_R9 <- "ER: no DIF"
          print( paste0("failure step5R contradicts step2; data: ", covars[i] ) )
        } else {
          print( paste0("failure step5R; M5.0R9 vs. M5.0; data: ", covars[i]) )
        }
      }
        
      }
      
    }
  }

         if ( exists("ERdif5_R9") ) { finMod <- paste0(ERdif5_R9, "; ", LFSdif4) 
  } else if ( exists("ERdif5_R8") ) { finMod <- paste0(ERdif5_R8, "; ", LFSdif4) # only structural DIF for ER
  } else if ( exists("ERdif5_R7") ) { finMod <- paste0(ERdif5_R7, "; ", LFSdif4) # non-structural DIF for ER 
  } else {                            finMod <- paste0(ERdif4,    "; ", LFSdif4) }

  bestModels <- rbind(bestModels, a1[ a1$nam == finMod, ] )
  ERdif5_R7 <- ERdif5_R8 <- ERdif5_R9 <- "" # create if it did not exists
  rm(list = c("ERdif2", "LFSdif2", "ERdif4", "LFSdif4", "ERdif5_R7", "ERdif5_R8", "ERdif5_R9"))
}

bestModels <- cbind(covars, bestModels) # add column with covariate names
names(bestModels)[1:2] <- c("covariate",  "description") # change column names
row.names(bestModels) <- NULL
bestModels
```

Save and clear.
```{r}
write.csv(x = bestModels,
          file = ifelse( test = resTes == "",
                         yes = paste0("final-LRT-2017.csv"),
                         no =  paste0("final-LRT-", resTes, "-2017.csv") ),
          row.names = FALSE)
rm(list = ls())
```

