---
title: "Exhaustive BIC method"
subtitle: "2017 dataset"
author: "Frederick Restrepo Estrada"
date: "2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


Made in **R version 4.1.3 (2022-03-10)**


Data used: `snapshot2017you.sav`

The dataset is a trimmed version (snapshot) of a dataset made available by CBS. It is not publicly available.


Use this template as follows:

1. In the chunk under "Define", indicate what type of restrictions are considered by commenting out all other lines with `resTes` (`ctrl+shift+C`). 

2. Run all chunks (`ctrl+alt+r`).


Function for extracting LG output.
```{r}
extract.output <- function(outPath, Nob) {

  output <- readLines(con = outPath) # read text lines
  output <- output[-1] # remove headers that are created when using `write` in LG
  output <- strsplit(x = output, split = ",") # LG creates `.csv` file
  maxLength <- max( sapply(X = output, FUN = length) ) # max number of columns in `.csv` file
  output <- lapply(X = output, FUN = `length<-`, maxLength) # create elements of equal length needed for dataframe
  output <- as.data.frame( do.call(rbind, output) ) # create dataframe
  row.names(output) <- NULL

  nams <- unique( output[ , 1] )
  note <- output[ which( output[ , 4] == 0 ), 5 ]
  note[is.na(note)] <- "OK"

  LLs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 6 ] ) # log-likelihoods
  # LPs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 8 ] ) # log-posteriors
  # L2s <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 9 ] ) # likelihood-ratio chi-square statistic
  # BVRs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 13 ] ) # total bivariate residuals
  Npars <- as.numeric( rowSums( !is.na(output[ which( output[, 5] == "Parameters" ), 6:ncol(output) ] ) ) )
  AIC.LL <- -2*LLs + Npars*2
  BIC.LL <- -2*LLs + Npars*log(Nob)
  # BIC.L2 <- L2s - df*log(Nob)

  return( data.frame(nam = nams, LL = LLs, Npars = Npars, AIC.LL, BIC.LL = BIC.LL, note = note) )
}
```

Load library.
```{r}
library(haven)
```


#### Exhaustive BIC method for a single covariate Z

Define.
```{r}
# resTes <- ""
# resTes <- "R7"
# resTes <- "R8"
resTes <- "R"

covars <- c(
  "age_group",
  "company_size",
  "economic_activity",
  "education_level",
  "interview_style",
  "job_length",
  "migration_background",
  "software_cluster",
  "work_hours"
  )

dat <- read_sav(file = "snapshot2017you.sav")
N <- nrow(dat)
```

Extract best model with BIC as evaluation criteria.
```{r}
bestModels <- data.frame()

for ( i in 1:length(covars) ) {
  
  outputName <- paste0("output-", covars[i], "-2017.csv") # name of table with LG output
  LGoutput <- extract.output( outPath = paste0(outputName), Nob = N)
  
  # remove restricted variants if unnecessary
  if (resTes == "") {
    LGoutput <- LGoutput[ !grepl(pattern = "(R7)", x = LGoutput$nam) &
                          !grepl(pattern = "(R8)", x = LGoutput$nam), ] # ignore all restrictions
  } else if (resTes == "R8") { 
    LGoutput <- LGoutput[ !grepl(pattern = "(R7)", x = LGoutput$nam), ] # ignore "Restriction #7"
  } else if (resTes == "R7") {
    LGoutput <- LGoutput[ !grepl(pattern = "(R8)", x = LGoutput$nam), ] # ignore "Restriction #8"
  }
  
  if ( any(LGoutput$note != "OK") ) { LGoutput$note <- paste0(LGoutput$note, "; SOME WARNING") }
  bestModels <- rbind( bestModels, LGoutput[which( LGoutput$BIC.LL == min(LGoutput$BIC.LL) ), ] )

}

bestModels <- cbind(covars, bestModels) # add column with covariate names
names(bestModels)[1:2] <- c("covariate",  "description") # change column names
row.names(bestModels) <- NULL
bestModels
```

Save.
```{r}
write.csv(x = bestModels,
          file = ifelse( test = resTes == "",
                         yes = paste0("final-BIC-2017.csv"),
                         no = paste0("final-BIC-", resTes, "-2017.csv") ),
          row.names = FALSE)
```

Clear environment.
```{r}
rm(list = ls())
```

