---
title: "Creating snapshot dataset"
author: "Frederick Restrepo Estrada"
date: "22-6-2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


The original datasets (`sav` files) with matched ER-LFS-data were provided by CBS. The trimmed datasets are derived from these datasets. Both the original datasets and the trimmed datasets are not publicly available.

Create a snapshot dataset as follows:

1. In the chunk under "Load data", comment out the lines for all but one year (`ctrl+shift+C`).

2. In the chunk under "Save snapshot datasets", comment out the lines for all but one age group.

3. Run all chunks (`ctrl+alt+r`).


Load libraries
```{r load_data}
library(haven)
library(data.table)
```


### Create snapshot dataset(s)

Load data
```{r}
# all LFS-cohors that started in 2016
dataYear <- read_spss("HMMdata2016_2017_voor_Frederick.sav")
year <- 2016

# all LFS-cohors that started in 2017
# dataYear <- read_spss("HMMdata2017_2018_voor_Frederick.sav")
# year <- 2017

# all LFS-cohors that started in 2018
# dataYear <- read_spss("HMMdata2018_2019_voor_Frederick.sav")
# year <- 2018

# transform to datatable
setDT(dataYear)
```

Inspect names
```{r}
names(dataYear)
```

Trim data
```{r}
# observations with records for both ER and LFS
dataYearB <- dataYear[!is.na(contract) & !is.na(contractEBB), ]

# first observation per respondent
dataYearB <- dataYearB[, .SD[1], by = persnr]

# 15-24 year-olds (young)
dataYearYou <- dataYearB[LftKlasse == 1 | LftKlasse == 2]

# 25-54 year-olds (middle-age)
dataYearMid <- dataYearB[LftKlasse == 3 | LftKlasse == 4]

# 55-74 year-olds (old)
dataYearOld <- dataYearB[LftKlasse == 5 | LftKlasse == 6]
```

Clean data
```{r}
# 15-24 year-olds
snapshotYearYou <- dataYearYou[, .(
  person_number = persnr,
  month_year = Maand,             # month with respect to January
  cohort_number = cohort,
  month_LFS = month0,             # month with respect to start LFS
  contract_ER = contract,
  # l1contract,
  # l3contract,
  contract_LFS = contractEBB,
  # l3contractEBB,
  # leeftijd,                     # numeric (detailed) version of "LftKlasse"
  age_group = LftKlasse,
  interview_style = proxy,
  gender = geslacht,
  migration_background = land,
  education_level = opleiding,
  economic_activity = SBIgroep,
  # SBIgroep2,                    # subset of "SBIgroep" with a larger "other" category level
  work_hours = Contracturen,
  # Contracturen2,                # subset of "Contracturen" with a larger "other" category level
  # BaanduurMaanden,              # numeric (detailed) version of "BaanduurMaanden"
  job_length = BaanduurKlasse,
  # BaanduurKlasse2,              # subset of "BaanduurKlasse" with a larger "other" category level
  software_cluster= softwarecluster,
  company_size = grootteklasse,
  same_ER = ZelfdebaanPolis,
  same_LFS = ZelfdewerkgevEBB
  # gewichtp1,
  # EBBGEWKWARTAALGEWICHTLONGA,
  # EBBGEWSTARTGEWP1,
  # EBBGEWJAARGEWICHTA,
  # EBBGEWKWARTAALGEWICHTLONG,
  # EBBGEWJAARGEWICHTLONGA,
  # EBBGEWJAARGEWICHTP1A,
  # EBBGEWKWARTAALGEWICHTA
  )]

# 25-54 year-olds
snapshotYearMid <- dataYearMid[, .(
  person_number = persnr,
  month_year = Maand,             # month with respect to Januari
  cohort_number = cohort,
  month_LFS = month0,             # month with respect to start LFS
  contract_ER = contract,
  # l1contract,
  # l3contract,
  contract_LFS = contractEBB,
  # l3contractEBB,
  # leeftijd,                     # numeric (detailed) version of "LftKlasse"
  age_group = LftKlasse,
  interview_style = proxy,
  gender = geslacht,
  migration_background = land,
  education_level = opleiding,
  economic_activity = SBIgroep,
  # SBIgroep2,                    # subset of "SBIgroep" with a larger "other" category level
  work_hours = Contracturen,
  # Contracturen2,                # subset of "Contracturen" with a larger "other" category level
  # BaanduurMaanden,              # numeric (detailed) version of "BaanduurMaanden"
  job_length = BaanduurKlasse,
  # BaanduurKlasse2,              # subset of "BaanduurKlasse" with a larger "other" category level
  software_cluster= softwarecluster,
  company_size = grootteklasse,
  same_ER = ZelfdebaanPolis,
  same_LFS = ZelfdewerkgevEBB
  # gewichtp1,
  # EBBGEWKWARTAALGEWICHTLONGA,
  # EBBGEWSTARTGEWP1,
  # EBBGEWJAARGEWICHTA,
  # EBBGEWKWARTAALGEWICHTLONG,
  # EBBGEWJAARGEWICHTLONGA,
  # EBBGEWJAARGEWICHTP1A,
  # EBBGEWKWARTAALGEWICHTA
  )]

# 55-74 year-olds
snapshotYearOld <- dataYearOld[, .(
  person_number = persnr,
  month_year = Maand,             # month with respect to Januari
  cohort_number = cohort,
  month_LFS = month0,             # month with respect to start LFS
  contract_ER = contract,
  # l1contract,
  # l3contract,
  contract_LFS = contractEBB,
  # l3contractEBB,
  # leeftijd,                     # numeric (detailed) version of "LftKlasse"
  age_group = LftKlasse,
  interview_style = proxy,
  gender = geslacht,
  migration_background = land,
  education_level = opleiding,
  economic_activity = SBIgroep,
  # SBIgroep2,                    # subset of "SBIgroep" with a larger "other" category level
  work_hours = Contracturen,
  # Contracturen2,                # subset of "Contracturen" with a larger "other" category level
  # BaanduurMaanden,              # numeric (detailed) version of "BaanduurMaanden"
  job_length = BaanduurKlasse,
  # BaanduurKlasse2,              # subset of "BaanduurKlasse" with a larger "other" category level
  software_cluster= softwarecluster,
  company_size = grootteklasse,
  same_ER = ZelfdebaanPolis,
  same_LFS = ZelfdewerkgevEBB
  # gewichtp1,
  # EBBGEWKWARTAALGEWICHTLONGA,
  # EBBGEWSTARTGEWP1,
  # EBBGEWJAARGEWICHTA,
  # EBBGEWKWARTAALGEWICHTLONG,
  # EBBGEWJAARGEWICHTLONGA,
  # EBBGEWJAARGEWICHTP1A,
  # EBBGEWKWARTAALGEWICHTA
  )]
```

Save snapshot datasets
```{r}
# 15-24 year-olds
write_sav( data = snapshotYearYou, path = paste0("snapshot", year, "you.sav") )

# 25-54 year-olds
# write_sav( data = snapshotYearMid, path = paste0("snapshot", year, "mid.sav") )

# 55-74 year-olds
# write_sav( data = snapshotYearOld, path = paste0("snapshot", year, "old.sav") )
```


### Extra information

First observations
```{r}
sum( is.na(dataYear[, .SD[1], by = persnr][, contract]) )
# ER observations are included at each time point

sum( is.na(dataYear[, .SD[1], by = persnr][, contractEBB]) )
# LFS observations are not included at each time point
```

Redundant variables
```{r}
all( attr(dataYear$SBIgroep2, "labels") %in% attr(dataYear$SBIgroep, "labels") )
# "SBIgroep2" does not contain unique information

all( attr(dataYear$Contracturen2, "labels") %in% attr(dataYear$Contracturen, "labels") )
# "Contracturen2" does not contain unique information

all( attr(dataYear$BaanduurKlasse2, "labels") %in% attr(dataYear$BaanduurKlasse, "labels") )
# "BaanduurKlasse2" does not contain unique information
```

Number of respondents per age group
```{r}
# 15-24 year-olds
nrow(dataYear[, .SD[1], by = persnr][LftKlasse == 1 | LftKlasse == 2])

# 25-54 year-olds
nrow(dataYear[, .SD[1], by = persnr][LftKlasse == 3 | LftKlasse == 4])

# 55-74 year-olds
nrow(dataYear[, .SD[1], by = persnr][LftKlasse == 5 | LftKlasse == 6])
```

