---
title: "Exhaustive AIC/BIC method"
subtitle: "Simulated datasets"
author: "Frederick Restrepo Estrada"
date: "2022"
output: html_document
---
 
```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


Made in **R version 4.1.3 (2022-03-10)**


Data used: `simDat.dat`, `simDat-step2.dat`


Use this template as follows:

1. In the chunk under "Define miscellaneous", indicate what criteria to use by commenting out `crit <- "AIC"` or `crit <- "BIC"` (`ctrl+shift+c`). 

2. Run all chunks (`ctrl+alt+r`).


Define working directory.
```{r}
pathR  <- "Y:/Frederick/latentGOLD/simulation3" # path to the folder that contains the LG output of all estimated models
```

Define functions.
```{r}
# Function for extracting LG output.
extract.output <- function(outPath, Nob) {
  
  output <- readLines(con = outPath) # read text lines
  output <- output[-1] # remove headers that are created when using `write` in LG
  output <- strsplit(x = output, split = ",") # LG creates `.csv` file
  maxLength <- max( sapply(X = output, FUN = length) ) # max number of columns in `.csv` file
  output <- lapply(X = output, FUN = `length<-`, maxLength) # create elements of equal length needed for dataframe
  output <- as.data.frame( do.call(rbind, output) ) # create dataframe
  row.names(output) <- NULL
  
  namRep <- strsplit(x = unique( output[ , 1] ), split = " | r", fixed = TRUE)
  nams <- sapply(X = namRep, FUN = "[[", 1)
  reps <- sapply(X = namRep, FUN = "[[", 2)
  note <- output[ which( output[ , 4] == 0 ), 5 ]
  note[is.na(note)] <- "OK"
  
  LLs <- as.numeric( output[ which( output[ , 4] == 1 ), 6 ] ) # log-likelihoods
  # LPs <- as.numeric( output[ which( output[ , 4] == 1 ), 8 ] ) # log-posteriors
  # L2s <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 9 ] ) # likelihood-ratio chi-square statistics
  # BVRs <- as.numeric( output[ which( output[ , 5] == "Statistics" ), 13 ] ) # total bivariate residuals
  Npars <- as.numeric( rowSums( !is.na(output[ which( output[, 4] == 5 ), 6:ncol(output) ] ) ) )
  AIC.LL <- -2*LLs + Npars*2
  BIC.LL <- -2*LLs + Npars*log(Nob)
  # BIC.L2 <- L2s - df*log(Nob)
  
  return( data.frame(nam = nams, rep = reps, LL = LLs, Npars = Npars, AIC.LL = AIC.LL, BIC.LL = BIC.LL, note = note) )
}
```


#### Exhaustive BIC method for a single covariate Z

Define miscellaneous.
```{r}
# criteria used
# crit <- "AIC"
crit <- "BIC"

claSep <- "s8" # class separation
numCat <- "c2" # number of categories

covEff <- rep( rep( rep( c("e10", "e11", "e20", "e21", "e22"), each = 10 ), each = 2 ), each  = 3 ) # covariate effect
effSiz <- rep( rep( rep( c("S", "M", "L"), times = 10 ), each = 2 ), times = 5 ) # effect size (odds-ratio)
casWei <- rep( rep( rep( c( 2000, 20000), each = 10), times = 3), times = 5)
repNum <- rep( rep( rep( 1:10, times = 2 ), times = 3), times = 5 )

covEff <- c( rep( "e00",          times = 20), covEff)
effSiz <- c( rep( "N",            times = 20), effSiz)
casWei <- c( rep( c(2000, 20000), each = 10 ), casWei )
repNum <- c( rep( 1:10,           times = 2), repNum )

s <- seq(from = 1, to = length(covEff), by = 10)

e <- data.frame(
  casWei = casWei[s], covEff = covEff[s], effSiz = effSiz[s],
  e00 = 0, e01 = 0, e02 = 0,
  e10 = 0, e11 = 0, e12 = 0,
  e20 = 0, e21 = 0, e22 = 0
)
```

Apply method.
```{r}
for ( i in 1:length(covEff) ) {
  outNam <- paste0("output-", claSep, "-", numCat, "-", covEff[i], "-", effSiz[i], "-w", casWei[i], "-r", repNum[i],  ".csv") # names of tables with LG output
  outExt <- extract.output( outPath = paste0(pathR, "/", outNam), Nob = casWei[i] )
  finMod <- outExt$nam[ outExt[ , paste0(crit, ".LL") ] == min( outExt[ , paste0(crit, ".LL") ] ) ]
  
  # save results in a data frame
  if (finMod == "y1: no; y2: no") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e00"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e00"] + 1
  } else if (finMod == "y1: uniform; y2: no") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e10"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e10"] + 1
  } else if (finMod == "y1: non-uniform; y2: no") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e20"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e20"] + 1
  } else if (finMod == "y1: no; y2: uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e01"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e01"] + 1
  } else if (finMod == "y1: uniform; y2: uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e11"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e11"] + 1
  } else if (finMod == "y1: non-uniform; y2: uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e21"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e21"] + 1
  } else if (finMod == "y1: no; y2: non-uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e02"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e02"] + 1
  } else if (finMod == "y1: uniform; y2: non-uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e12"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e12"] + 1
  } else if (finMod == "y1: non-uniform; y2: non-uniform") {
    e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e22"] <-
      e[e$covEff == covEff[i] & e$effSiz == effSiz[i] & e$casWei == casWei[i], "e22"] + 1
  }
  
}
```

Save results.
```{r}
e <- e[ with( data = e, expr = order(casWei, covEff) ), ]
write.csv(x = e, file = paste0("final-", crit, "-", claSep, "-", numCat, ".csv"), row.names = FALSE)
```

