---
title: "Simulation covariates"
subtitle: "Parameters and stepwise selection"
author: "Frederick Restrepo Estrada"
date: "2022"
output: html_document
---

Load packages.
```{r}
library(reshape)
library(nnet)
```

Create dataset with results from all methods.
```{r}
#######################
## fetch LRT results ##
#######################

# read data (in wide format as a frequency table)
dLRT <- read.csv(file = "final-LRT-s8-c2.csv", header = TRUE, sep = ",")

# convert to long format (frequency tale) 
dLRT <- melt( data = dLRT, id.vars = c("casWei", "covEff", "effSiz") )

# keep only entries for which a "final" model was found
dLRT <- dLRT[dLRT$value != 0, ]

# initialize case format dataframe
cLRT <- setNames( object = data.frame( matrix(data = NA, nrow = 0, ncol = ncol(dLRT) ) ), nm = names(dLRT) ) 
# fill in case format dataframe
for ( i in 1:nrow(dLRT) ) {
  # for ( i in 1:5 ) {
  v <- as.numeric( dLRT[i, "value"] )
  for ( j in 1:v ) {
    cLRT <- rbind(cLRT,  dLRT[i, ])
  }
}

# add column
cLRT$result <- factor( ifelse(test = cLRT$variable == cLRT$covEf,
                              yes = "all", no = ifelse(test = cLRT$variable != "e00",
                                                       yes = "some", no = "none") ) )

# remove unnecessary column
cLRT$value <- NULL

# remove "no effect" category
cLRT <- cLRT[cLRT$effSiz != "N", ]

# set variable class
cLRT$casWei <- as.factor(cLRT$casWei)
cLRT$covEff <- as.factor(cLRT$covEff)
cLRT$effSiz <- as.factor(cLRT$effSiz)
cLRT$variable <- as.factor(cLRT$variable)

# reorder dataframe
cLRT <- cLRT[order(cLRT$casWei, cLRT$covEff, cLRT$effSiz, decreasing = c(FALSE, FALSE, TRUE) ), ]

#######################
## fetch BIC results ##
#######################

dBIC <- read.csv(file = "final-BIC-s8-c2.csv", header = TRUE, sep = ",")
dBIC <- melt(data = dBIC, id.vars = c("casWei", "covEff", "effSiz"))
dBIC <- dBIC[dBIC$value != 0, ]

cBIC <- setNames( object = data.frame( matrix(data = NA, nrow = 0, ncol = ncol(dBIC) ) ), nm = names(dBIC) )
for ( i in 1:nrow(dBIC) ) {
  v <- as.numeric( dBIC[i, "value"] )
  for ( j in 1:v ) {
    cBIC <- rbind(cBIC,  dBIC[i, ])
  }
}
row.names(cBIC) <- NULL
cBIC$result <- factor( ifelse(test = cBIC$covEf == cBIC$variable,
                              yes = "all", no = ifelse(test = cBIC$variable != "e00",
                                                       yes = "some", no = "none") ) )
cBIC <- cBIC[cBIC$effSiz != "N", ]
cBIC$value <- NULL

cBIC$casWei <- as.factor(cBIC$casWei)
cBIC$covEff <- as.factor(cBIC$covEff)
cBIC$effSiz <- as.factor(cBIC$effSiz)
cBIC$variable <- as.factor(cBIC$variable)

cBIC <- cBIC[order(cBIC$casWei, cBIC$covEff, cBIC$effSiz, decreasing = c(FALSE, FALSE, TRUE) ), ]

#######################
## fetch AIC results ##
#######################

dAIC <- read.csv(file = "final-AIC-s8-c2.csv", header = TRUE, sep = ",")
dAIC <- melt( data = dAIC, id.vars = c("casWei", "covEff", "effSiz") )
dAIC <- dAIC[dAIC$value != 0, ]

cAIC <- setNames( object = data.frame( matrix(data = NA, nrow = 0, ncol = ncol(dAIC) ) ), nm = names(dAIC) )
for ( i in 1:nrow(dAIC) ) {
  v <- as.numeric( dAIC[i, "value"] )
  for ( j in 1:v ) {
    cAIC <- rbind(cAIC,  dAIC[i, ])
  }
}
row.names(cAIC) <- NULL
cAIC$result <- factor( ifelse(test = cAIC$covEf == cAIC$variable,
                              yes = "all", no = ifelse(test = cAIC$variable != "e00",
                                                       yes = "some", no = "none") ) )
cAIC <- cAIC[cAIC$effSiz != "N", ]
cAIC$value <- NULL

cAIC$casWei <- as.factor(cAIC$casWei)
cAIC$covEff <- as.factor(cAIC$covEff)
cAIC$effSiz <- as.factor(cAIC$effSiz)
cAIC$variable <- as.factor(cAIC$variable)

cAIC <- cAIC[order(cAIC$casWei, cAIC$covEff, cAIC$effSiz, decreasing = c(FALSE, FALSE, TRUE) ), ]

######################
## combine datasets ##
######################

# add methods
cLRT$method <- "LRT"
cBIC$method <- "BIC"
cAIC$method <- "AIC"

# combine datsets
dat <- rbind(cLRT, cBIC, cAIC)
dat$method <- as.factor(dat$method)
dat$effSiz <- factor( x = dat$effSiz, levels = c("S", "M", "L") )
dat$method <- factor( x = dat$method, levels = c("LRT", "AIC", "BIC") )
dat$result <- factor( x = dat$result, levels = c("none", "some", "all") )
dat <- dat[order(dat$casWei, dat$covEff, dat$effSiz, dat$method, decreasing = c(FALSE, FALSE, TRUE, TRUE) ), ]
row.names(dat) <- NULL

# z value = estimate / std. error
```


#### Parameters

Coefficients, standard errors and p-values.
```{r}
MLR1b <- multinom(formula = result ~ casWei + covEff + effSiz + method,
                  data = dat, trace = FALSE)

b <- summary(MLR1b)

###################
## coefficients  ##
###################

print("ceof MLR")
b$coefficients

#########
## SEs ##
#########

print("SEs MLR")
b$standard.errors

###########
## pvals ##
###########

zValMLR <- b$coefficients / b$standard.errors

print("pvals MLR")
2 * pnorm(q = abs(x = zValMLR), mean = 0, sd = 1, lower.tail = FALSE)
```


#### Stepwise selection

Stepwise selection can be performed with the function `step`. The function allows for the selection criteria to be changed by changing `k`. 

Construct model.
```{r}
MLR1b <- multinom(formula = result ~ casWei + covEff + effSiz + method,
                  data = dat, trace = FALSE)
```

Test.
```{r}
# with AIC
step( object = MLR1b, direction = "backward", k = 2)

# with BIC
step( object = MLR1b, direction = "backward", k = log( nrow(dat) ) )

# AIC full model
MLR1b$edf * 2 + MLR1b$deviance
```

The function `step` seems to ignore `k` when given a `multinom` object. `step` seems to use AIC only. Therefore, stepwise selection with BIC is performed manually.

BIC for `multinom` can be calculated with:
- `MLR1b$edf * log( nrow(dat) ) + MLR1b$deviance`

AIC for `multinom` can be calculated with:
- `MLR1b$edf * 2 + MLR1b$deviance`

Note: residual deviance = -2 * LL

Backwards selection using AIC.
```{r}
##################
## BACKWARD AIC ##
##################

MLR1b <- multinom(formula = result ~ casWei + covEff + effSiz + method,
                  data = dat, trace = FALSE)

step(object = MLR1b, direction = "backward",
     scope = (~ casWei + covEff + effSiz + method), k = 2 )
```

Backward selection using BIC.
```{r}
##################
## BACKWARD BIC ##
##################

# none
tmp <- multinom(formula = result ~ casWei + covEff + effSiz + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# - casWei
tmp <- multinom(formula = result ~ covEff + effSiz + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# - covEff
tmp <- multinom(formula = result ~ casWei + effSiz + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# - effSiz
tmp <- multinom(formula = result ~ casWei + covEff + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# - method
tmp <- multinom(formula = result ~ casWei + covEff + effSiz,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance
```

With 'backwards selection' using BIC, the final model keeps all covariates.

Forward selection using AIC.
```{r}
#################
## FORWARD AIC ##
#################

MLR0b <- multinom(formula = result ~ 1,
                  data = dat, trace = FALSE)

step(object = MLR0b, direction = "forward",
     scope = (~ casWei + covEff + effSiz + method), k = 2 )
```

Forward selection using BIC.
```{r}
#################
## FORWARD BIC ##
#################

# step1; none
tmp <- multinom(formula = result ~ 1,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ casWei,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ effSiz,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# step2; + covEff
tmp <- multinom(formula = result ~ covEff,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + casWei,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + effSiz,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# step3; + effSiz
tmp <- multinom(formula = result ~ covEff + effSiz,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + effSiz + casWei,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + effSiz + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# step4; + method
tmp <- multinom(formula = result ~ covEff + effSiz + method,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

tmp <- multinom(formula = result ~ covEff + effSiz + method + casWei,
                data = dat, trace = FALSE)
tmp$edf * log(nrow(dat)) + tmp$deviance

# + casWei
```

With 'forward selection' using BIC, the final model keeps all covariates

