---
title: "Generate LG syntax"
subtitle: "2018 dataset, all models"
author: "Frederick Restrepo Estrada"
date: "2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


Made in **R version 4.1.3 (2022-03-10)**


Data used: `snapshot2018you.sav`

The dataset is a trimmed version (snapshot) of a dataset made available by CBS. It is not publicly available.


Create LG syntax as follows:

1. In the chunk under "Define covariate to be tested", comment out the lines for all but covariate (`ctrl+shift+C`).

2. Run all chunks (`ctrl+alt+r`).


#### Creating all possible models for a single covariate

Load libraries.
```{r}
library(brew)
library(haven)
```

Define covariate to be tested.
```{r}
# covar <- "age_group"
# covar <- "company_size"         # unknown / not applicable
# covar <- "economic_activity"    # unknown / not applicable
# covar <- "education_level"      # unknown
# covar <- "interview_style"
# covar <- "job_length"           # unknown / not applicable
# covar <- "migration_background"
# covar <- "software_cluster"     # unknown / not applicable
covar <- "work_hours"           # unknown / not applicable
```

Set default values.
```{r}
dat <- read_sav(file = "snapshot2018you.sav") # load dataset
N <- nrow(dat) # number of cases
nL <- 3 # number of latent classes

nI <- nrow( unique( dat[, "contract_ER"] ) ) # number of indicator categories
nC <- nrow( unique( dat[ , covar] ) ) # number of covariate categories

#################
## MAIN EFFECT ##
#################

# total number of "main" parameters
num_M <- (nI-1)*(nC-1)

# "main" parameters relating to derived frequencies
# Restriction #7 (R7): parameters set to zero that relate to derived frequencies
pM7 <- sort( 
  seq(from = num_M-(nI-1)+1, to = num_M, by = 1) # "right column" LG index
)

# "main" parameters relating to non-derived frequencies
# Restriction #8 (R8): parameters set to zero that DO NOT relate to derived frequencies
pM8 <- setdiff( 1:num_M, pM7 )

########################
## INTERACTION EFFECT ##
########################

# total number of "interaction" parameters
num_I <- (nI-1)*(nC-1)*(nL-1)

# "interaction" parameters relating to derived frequencies
# Restriction #7 (R7): parameters set to zero that relate to derived frequencies
pI7 <- sort(
  seq(from = num_I-(nI-1)*(nL-1)+1, to = num_I, by = 1) # "right column" LG index
)

# "interaction" parameters relating to non-derived frequencies
# Restriction #8 (R8): parameters set to zero that DO NOT relate to derived frequencies
pI8 <- setdiff( 1:num_I, pI7 )
```

Set default values (continued).
```{r}
restrict <- ifelse( test = covar %in% c("company_size", "economic_activity", "education_level", "job_length", "software_cluster", "work_hours"), yes = "with", no = "without" ) # if the last category is "unknown" or "unknown / not applicable"

if ( restrict == "with" ) {
  z <- expand.grid( ER = c( "no DIF",
                            "uniform DIF (R7)", "uniform DIF (R8)", "uniform DIF",
                            "non-uniform DIF (R7)", "non-uniform DIF (R8)", "non-uniform DIF" ),
                    LFS = c("no DIF", "uniform DIF", "non-uniform DIF") )
} else if ( restrict == "without" ) {
  z <- expand.grid( ER = c( "no DIF", "uniform DIF", "non-uniform DIF"),
                    LFS = c( "no DIF", "uniform DIF", "non-uniform DIF") ) 
}
```

Create LG syntax file. 
```{r}
for ( i in 1:nrow(z) ) {
  
  modNam <- paste0("ER: ", z[i,1], "; LFS: ", z[i,2])
  outToDo <- paste0( ifelse(test = i == 1, yes = "write=", no = "append="),
                     paste0("'output-" , covar, "-2018.csv'") )
  indepVar <- paste(",", covar, "nominal")
  indepEqu <- paste("+", covar)
  
  ###################
  ## 1st indicator ##
  ###################
  if ( z[i,1] == "no DIF" ) {
    aDIF <- aPar <- ""
  } else if (z[i,1] == "uniform DIF (R7)") {
    aDIF <- paste0("+ (a1) ", covar)
    aPar <- paste0("a1[1,", pM7, "] = 0;", collapse = " ")
  } else if (z[i,1] == "uniform DIF (R8)") {
    aDIF <- paste0("+ (a1) ", covar)
    aPar <- paste0("a1[1,", pM8, "] = 0;", collapse = " ")
  } else if (z[i,1] == "uniform DIF") {
    aDIF <- paste0("+ (a1) ", covar)
    aPar <- ""
  } else if (z[i,1] == "non-uniform DIF (R7)") {
    aDIF <- paste0("+ (a1) ", covar, " + (a2) cluster ", covar)
    aPar <- paste( c( paste0("a1[1,", pM7, "] = 0;"),                    # main parameters
                      paste0("a2[1,", pI7, "] = 0;") ), collapse = " " ) # interaction parameters
  } else if (z[i,1] == "non-uniform DIF (R8)") {
    aDIF <- paste0("+ (a1) ", covar, " + (a2) cluster ", covar)
    aPar <- paste( c( paste0("a1[1,", pM8, "] = 0;"),
                      paste0("a2[1,", pI8, "] = 0;") ), collapse = " " )
  } else if (z[i,1] == "non-uniform DIF") {
    aDIF <- paste0("+ (a1) ", covar, " + (a2) cluster ", covar)
    aPar <- ""
  }
  
  ###################
  ## 2nd indicator ##
  ###################
  if ( z[i,2] == "no DIF" ) {
    bDIF <- bPar <- ""
  } else if (z[i,2] == "uniform DIF") {
    bDIF <- paste0("+ (b1) ", covar)
    bPar <- ""
  } else if (z[i,2] == "non-uniform DIF") {
    bDIF <- paste0("+ (b1) ", covar, " + (b2) cluster ", covar)
    bPar <- ""
  }
  
  ######################
  ## assign variables ##
  ######################
  temEnv <- new.env()
  
  assign(x = "datPat", value = "snapshot2018you.sav")
  
  assign(x = "modNam",   value = modNam,   envir = temEnv)
  assign(x = "outToDo",  value = outToDo,  envir = temEnv)
  assign(x = "indepVar", value = indepVar, envir = temEnv)
  assign(x = "indepEqu", value = indepEqu, envir = temEnv)
  assign(x = "aDIF",     value = aDIF,     envir = temEnv)
  assign(x = "bDIF",     value = bDIF,     envir = temEnv)
  assign(x = "aPar",     value = aPar,     envir = temEnv)
  assign(x = "bPar",     value = bPar,     envir = temEnv)
  
  ########################
  ## create syntax file ##
  ########################
  if (i == 1) {
    synNam1 <- paste0("syntax-model-", covar, "-2018.lgs")
    template <- file(description = "template-syntax-file.brew", open = "r")
    brew(template, output = synNam1, envir = temEnv)
    close(con = template)
  }
  
  synNam2 <- "syntax-part.lgs"
  template <- file(description = "template-syntax-model.brew", open = "r")
  brew(template, output = synNam2, envir = temEnv)
  close(con = template)
  
  file.append(file1 = synNam1, file2 = synNam2)
}
```

Remove variables.
```{r}
rm( list = ls() )
```

